---
title: "Data 606 - Assignment 12"
author: "Joseph Simone"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Regression

The attached who.csv dataset contains real-world data from 2008. The variables included follow.
Country: name of the country

LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
TotExp: sum of personal and government expenditures.

```{r}
who <- read.csv("who.csv")
head(who)
```

```{r}
summary(who)
```


### 1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}
ggplot(data = who, aes(x=who$LifeExp, y=who$TotExp)) + geom_point(size = 2, alpha = .4) + geom_smooth(method = "lm", se = FALSE)
```

```{r}
linearmodel =lm(LifeExp~ TotExp, data = who)

summary(linearmodel)
```

The linear model is as follow: LifeExp=6.475e+6.297e-05*TotExp

It is a a statistically significant predictor of evaluation score with p-value less than 0.05. But for Multiple R-squared and R-squared, the model is only around 25% fits the data.

```{r}
qqnorm(linearmodel)$residuals)
qqline(linearmodel)$residuals)
```

Especially have deviation at lower and quantiles. The residuals does not show randomly.

### 2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

```{r}
raiseLifeExp <- who$LifeExp^4.6
raiseTotExp <- who$TotExp^0.06
```

```{r}
linearmodel2 =lm(raiseLifeExp~ raiseTotExp, data = who)

summary(linearmodel2)
```

the linear model is as follow: LifeExp=-736527910+620060216*TotExp

It is a a statistically significant predictor of evaluation score with p-value less than 0.05. For Multiple R-squared and R-squared, the model is only around 72% fits the data.

This model using the transformed variables has a larger Multiple R-squared and R-squared, so it is better than first one.

```{r}
qqnorm(linearmodel2$residuals)
qqline(linearmodel2$residuals
```

Q-Q plot are more uniformly scattered than first oneand still have deviation at lower and quantiles.

### 3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5. 

$LifeExp=((-736527910+620060216)(TotExp))$


```{r}
TotExp <- 1.5
raiseLife <- -736527910 + 620060216*TotExp
raiseLife
```

```{r}
LifeExp <- raiseLife ^(1/4.6)
LifeExp
```

forecast life expectancy is 63.3 when TotExp^.06 =1.5.

#### Forest lift expectancy?
When $TotExp^.06 =2.5$
```{r}
TotExp <- 2.5

raiseLife
```

```{r}
LifeExp <- raiseLife ^(1/4.6)
LifeExp
```

Forecast life expectancy is 86.5 when $TotExp^.06 =2.5$.

### 4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?

$LifeExp = ((b0+b1)(PropMd))+ ((b2)(TotExp)) +(((b3)(PropMD)(TotExp)))$

```{r}
mmodel <- lm(LifeExp ~ PropMD + TotExp + PropMD*TotExp, data = who)
summary(mmodel)
```


the linear model is as follow: LifeExp=6.27710 + 1.497103PropMD + 7.23310-5TotExp -6.02610^-3 PropMDTotExp

F-statistic with P-value < 0.05, it is a a statistically significant predictor of evaluation score. The Multiple R-squared and adjusted R-squared is only 34%-35% fit the data.

The F-Statistic of 34.49 with a very small p-value show us that the currect model is better than the model with one less predictor. The R2R2 value of .357 means the model explains 35.7% of the data which isn’t very good. The Adjusted R2R2 is more accurate by filtering out the noise but with a slightly lower value of .347. The p-values are still very small meaning all the variables are very relevant to the model. The Std. Errors are good for the intercept, PropMD, TotExp but not for other (TotExp X PropMD). The Residuals look like they have a normal distribution around zero which is good.

```{r}
qqnorm(mmodel$residuals)
qqline(mmodel$residuals)
```


### 5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

```{r}
PropMD <- 0.03
TotExp <- 14
LifeExp<- (6.277*10) + (1.497*10^3)*PropMD + (7.233*10^-5)*TotExp - (6.026*10^-3) *PropMD*TotExp
LifeExp
```

A Life Expectancy of 107 is not plausible today.